---
title: "Interview_Upgrad"
author: "Milan A Joshi"
date: "16 October 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading data Sets and knowing structure
```{r}
basedt <-read.csv("baseball_train.csv")

baseball_unknown<-read.csv("baseball_unknown.csv")

str(basedt)
```
#Coverting Playoffs(Response) to factor
```{r}
basedt$Playoffs<-as.factor(basedt$Playoffs)

str(basedt)
```
#Removing unnecessary columns and calcuting summary statistics
```{r}
basedt<-basedt[,-c(1,2)]

summary(basedt)
```
#Loading caret package for preprocessing scaling and centering
```{r}

library(caret)

dt<-as.data.frame(basedt[,c(1,2)])

prpc<-preProcess(dt, method = c("center", "scale", "YeoJohnson"))

dt<-predict(prpc, newdata = dt)

basedt<-cbind(dt,basedt[,-c(1,2)])

summary(basedt)

str(basedt)

```
## Exploratory Analysis
# Analyzing Missing value in data
```{r}

Missingvalues<-apply(basedt,2,function(x) sum(is.na(x)))

Missingvalues

```
#Visualizing Missing value pattern with library VIM
```{r echo=FALSE}
library(VIM)

aggr_plot <- aggr(basedt, col=c('navyblue','red'), numbers=TRUE,     
                  
                  sortVars=TRUE, labels=names(basedt), cex.axis=.7,   
                  
                  gap=3, ylab=c("Histogram of missing data","Pattern"))
```
# Proportion of Playoffs 0-1 with ggplot2
```{r echo=FALSE}
library(ggplot2)

library(scales)

ggplot(basedt, aes(x = as.factor(Playoffs))) +
  
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  
  geom_text(aes(y = ((..count..)/sum(..count..)), 
                
                label = scales::percent((..count..)/sum(..count..))), 
            
                stat = "count", vjust = -0.25) +   
           
                 scale_y_continuous(labels = percent) +
  
  labs(title = "Playoffs 0 vs 1", y = "Percent", x = "Playoffs")

```
# Analysing Numeric Varibable using correlations and Densities        AppliedpredictiveModeling Library

```{r echo=FALSE,warning=FALSE}

library(AppliedPredictiveModeling)

transparentTheme(trans = .5)


featurePlot(x = basedt[, 1:5],y = basedt$Playoffs,plot = "pairs",
            
            ## Add a key at the top
            
            auto.key = list(columns = 2))


transparentTheme(trans = .9)

featurePlot(x = basedt[, 1:5], 
            
            y = basedt$Playoffs,
            
            plot = "density", 
            
            ## Pass in options to xyplot() to
            
            ## make it prettier
            
            scales = list(x = list(relation="free"), 
                          
                          y = list(relation="free")), 
            adjust = 1.5, 
            
            pch = "|", 
            
            layout = c(5, 1), 
            
            auto.key = list(columns = 2))

featurePlot(x = basedt[, 1:5], 
            
            y = basedt$Playoffs, 
            
            plot = "box", 
            
            ## Pass in options to bwplot() 
            
            scales = list(y = list(relation="free"),
                  
                                  x = list(rot = 90)),  
            layout = c(5,1 ), 
            
            auto.key = list(columns = 2))
```
#Plotting correlations using PerformanceAnalytics with Densities
```{r echo=FALSE}

library(PerformanceAnalytics)

library(corrplot)

library(Hmisc)

my_data <- basedt[, c(1,2,3,4,5)]

chart.Correlation(my_data, histogram=TRUE, pch=19)

corrln <- cor(basedt[,1:5])

my_dt<-as.matrix(round(corrln, 2))

corrplot(my_dt, type = "upper", order = "hclust",tl.col = "black", 
         
           tl.srt = 45)

```
# Data Spliting 70-30 train-test
```{r}
set.seed(998)

inTraining <- createDataPartition(basedt$Playoffs, p = .70, list = FALSE)

train <- basedt[ inTraining,]

test <- basedt[-inTraining,]

train$Playoffs<-as.factor(train$Playoffs)

test$Playoffs<-as.factor(test$Playoffs)

```
# Model tuning 10 fold crossvalidation tuning parameters
```{r}

fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           ## Estimate class probabilities
                           classProbs = TRUE,
                           ## Evaluate performance using 
                           ## the following function
                           summaryFunction = twoClassSummary)

```
## Training gbm randomForest extraTree Logistic Regression svm(Kernel)
# Training gbm
```{r echo = FALSE}

set.seed(1001)

gbmFit<- train(Playoffs ~ ., data = train, 
                    method = "gbm", 
                    trControl = fitControl,
                    ## This last option is actually one
                    ## for gbm() that passes through
                    verbose = FALSE)
gbmFit

trellis.par.set(caretTheme())

plot(gbmFit)

```
# Performance of gbm on tset set
```{r}
test$pred <- predict(gbmFit, newdata =test)

confusionMatrix(data = test$pred, reference = test$Playoffs)
```
# Training RandomForest 
```{r }
set.seed(1002)

rfFit<- train(Playoffs~ ., data = train, 
                    method = "rf", 
                    trControl = fitControl, 
                    #preProc = c("center", "scale"),
                    tuneLength = 8,
                    metric = "ROC")
rfFit

trellis.par.set(caretTheme())

plot(rfFit)
```
# Predicting RandomForest on test Data 
```{r}

test$pred <- predict(rfFit, newdata =test)

confusionMatrix(data = test$pred, reference = test$Playoffs)

```
# Model Tuning for extraTree
```{r}
set.seed(997)

cctrl<- trainControl(method = "cv", number = 10, repeats=10,
                     
                     returnResamp = "all")
```
# Training extraTree Model
```{r echo= FALSE}
set.seed(1003)

extraTreeFit<- train(Playoffs~.,data=train ,
                        method = "extraTrees", 
                        trControl = cctrl,
                        #preProc = c("center", "scale"),
                        ntree = 50)

extraTreeFit

trellis.par.set(caretTheme())

plot(extraTreeFit)
```
# Performance extraTree and Predction on test set
```{r }
test$pred <- predict(extraTreeFit, newdata =test)

confusionMatrix(data = test$pred, reference = test$Playoffs)
```
# Training Logistic Regression
```{r }
set.seed(1004)

logitFit<- train(Playoffs ~. 
                    ,data=train, method="glm", 
                       trControl = fitControl,family="binomial")

logitFit

summary(logitFit)
```
# Performance Logistic Regression and prediction on test set
```{r }

test$pred <- predict(logitFit, newdata =test)

confusionMatrix(data = test$pred, reference = test$Playoffs)
```
# Training Support vector machine with gaussian kernel
```{r echo=FALSE}

set.seed(1005)

svmFit<-train(Playoffs~., data = train, 
              method = "svmRadial", 
              trControl = fitControl, 
              #preProc = c("center", "scale"),
              #tuneLength = 8,
              metric = "ROC")
svmFit

test$pred <- predict(svmFit, newdata =test)

confusionMatrix(data = test$pred, reference = test$Playoffs) 

svmFit

trellis.par.set(caretTheme())

plot(extraTreeFit)

```
# Performance and prediction of svm on test set
```{r }

test$pred <- predict(svmFit, newdata =test)

confusionMatrix(data = test$pred, reference = test$Playoffs) 
```
# Model Comaparison
```{r echo=FALSE}
resamps <- resamples(list(GBM = gbmFit,
                          SVM = svmFit,
                          RDF = rfFit,
                          #ExTr=extraTree,
                          LogisTicReg=logitFit))
                          
resamps

trellis.par.set(theme)

bwplot(resamps, layout = c(3, 1))

trellis.par.set(caretTheme())

dotplot(resamps, metric = "ROC")

splom(resamps)
```
# Plot of difference between Accuracy Metrics
```{r echo= FALSE}

difValues <- diff(resamps)

difValues

summary(difValues)

trellis.par.set(theme)

bwplot(difValues, layout = c(3, 1))

trellis.par.set(caretTheme())

dotplot(difValues)
```
#variable Importance in each model
```{r echo = FALSE}

gbmImp <- varImp(gbmFit, scale = FALSE)

gbmImp

plot(gbmImp, top = 3)

rfImp <- varImp(rfFit, scale = FALSE)

rfImp

plot(rfImp, top = 3)

ETImp <- varImp(extraTreeFit, scale = FALSE)

ETImp

plot(ETImp, top = 3)

logitImp <- varImp(logitFit, scale = FALSE)

logitImp

plot(logitImp, top = 3)

svmImp <- varImp(svmFit, scale = FALSE)

svmImp

plot(svmImp, top = 20)
```
# Most Important variables using simulatedAnneling
```{r echo=FALSE}
sa_ctrl <- safsControl(functions = rfSA,
                       method = "repeatedcv",
                       repeats = 5,
                       improve = 50)

set.seed(10)

rf_sa <- safs(x = train[,-6], y =train$Playoffs,
              iters = 200,
              safsControl = sa_ctrl)
rf_sa

library(ggplot2)

plot(rf_sa) + theme_bw()
```

